A Fast kNN Algorithm Using Multiple Space-Filling Curves

### 2022-04-13 2022/Sensors/kNN.tex
fix formulas numbering

Let $\omega_i, 1 \leq i \leq S$, be the $i$-th object from the set of objects correlated with class $C_j$. Previously, a feature vector $y = (y_1, y_2, ..., y_N) \in R^N$, corresponding to each object under consideration was defined, i.e. 
\begin{equation}
\omega_i \leftrightarrow y = (y_{i1}, y_{i2}, ..., y_{iN}).
\end{equation}

Using the Peano curve, each object is assigned the value of $x_i$, i.e.
\begin{equation}
\omega_i \leftrightarrow x_i, \; y(x_i) = (y_{i1}, y_{i2}, ..., y_{iN}).
\end{equation}

Consider a test object $\omega^*$, to which the value $x^*$ is assigned. The degree of proximity of the test object $\omega^*$ to one of the objects $\omega_i$ is determined based on the distance on the one-dimensional scale $d_i = |x_i - x^* |$.
Thus, the closest object to the test object has the following property:
\begin{equation}
d_{min} = \min \{ d_i: 1 \leq i \leq S \}.
\end{equation}

The operation of the algorithm to determine the nearest neighbour of the test object involves two phases: data preparation and the search itself. Data preparation consists of calculating the correspondence $\omega_i \rightarrow x_i$ for each input object, including the test object, and ordering the objects by the value of $x$.
\begin{equation}
x^1(\omega^1) \leq x^2(\omega^2) \leq ... \leq x^S(\omega^S).
\end{equation}

@@

It follows from formulae (\ref{hypercubeD0}) and (\ref{maps}) that if an interval $d(M,v)$ is mapped onto the subcube $D_0(M,v)$, then there exists a family of subcubes
\begin{equation}
D_l(M,v) = y^l(d(M,v)), 1 \leq l \leq L,
\end{equation} 
connected with the corresponding subintervals $d_l(M,v_l) \subset [0,1]$, where $d(M,v) = d_0(M,v_0), v_0 = v$, such that 
\begin{equation}
D_l(M,v) = y^l(d_l(M,v)), 1 \leq l \leq L.
\end{equation}

Since the hypercube $D$ from (\ref{D}) belongs to the common part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Fig.~\ref{fig2}), then, by introducing an additional function 
\begin{equation}
g_0(y) = \max \{ |y_i|-2^{-1}: \; 1 \leq i \leq N \},
\end{equation} 
the original hypercube $D$ can be represented as
\begin{equation}
D = \{y^l(x): x \in [0,1], \; g_0(y^l(x)) \leq 0 \}, \; 0 \leq l \leq L,
\end{equation} 
i.e. $g_0(y) \leq 0$, if $y \in D$, and $g_0(y) > 0$ if otherwise. Hence, any point $y \in D$ has its preimage $x^l \in [0,1]$ at every mapping $y^l(x), \; 0 \leq l \leq L$, i.e.
\begin{equation}
y = y^l(x^l), \; x^l \in [0, 1], \; 0 \leq l \leq L.
\end{equation}

@@

\textbf{Theorem 1.} Let $y^*$ be an arbitrary point from the domain $D$ belonging to the interval with endpoints $y^\prime, y^{\prime\prime} \in D$, which differ in the values of the only coordinate, and let
\begin{equation}
|y_j^\prime - y^{\prime\prime}_j| \leq 2^{-p}, \; y^\prime_i = y^{\prime\prime}_i = y_i^*, \; 1 \leq i \leq N, i \ne j,
\end{equation} 
where $p$ is an integer, $1 \leq p \leq L-1$, and $j$ is the number of coordinate whose values for points $y^*, y^\prime, y^{\prime\prime}$ are different. Then there exists at least one correspondence $y^l(x), 0 \leq l \leq L$, and preimages $x^*, x^\prime, x^{\prime\prime} \in [0, 1]$ such that
\begin{equation}
 y^* = y^l(x^*), y^\prime = y^l(x^\prime), y^{\prime\prime} = y^l(x^{\prime\prime}), \\
\max \{ |x^\prime - x^*|, |x^{\prime\prime} - x^*|, |x^\prime - x^{\prime\prime}| \} \leq 2^{-pN}.
\end{equation}

@@

The scheme of implementation of the kNN algorithm using multiple Peano curves will look as follows.
For every object $\omega_i$ a set of its preimages on different one-dimensional scales is assigned 
\begin{equation}
\omega_i \rightarrow \{ x_i^0, x_i^1, ..., x_i^L \},
\end{equation} 
where $l, \; 0 \leq l \leq L$, is the number of the one-dimensional interval.

The degree of proximity of a test object $\omega^*$ to one of the known objects $\omega_i$ can be expressed in terms of the distance
\begin{equation}
d_i^l = |x_i^l - x^{*l}|
\end{equation} 
on the $l$-th scale.

Thus, the distance $d(\omega_i,\omega^*)$ between two objects $\omega_i$ and $\omega^*$ on a one-dimensional scale will correspond to the minimum distance among all the mappings 
\[
d(\omega_i,\omega^*) = \min_{0\leq l\leq L} d_i^l = \min_{0\leq l\leq L}  |x_i^l - x^{*l}| .
\]
If we find the minimum distance
\begin{equation}
d_{min} = \min_{1\leq i\leq n} d(\omega_i,\omega^*) ,
\end{equation} 
we get the nearest neighbour.

>>>>

Let $\omega_i, 1 \leq i \leq S$, be the $i$-th object from the set of objects correlated with class $C_j$. Previously, a feature vector $y = (y_1, y_2, ..., y_N) \in R^N$, corresponding to each object under consideration was defined, i.e. 
\[
\omega_i \leftrightarrow y = (y_{i1}, y_{i2}, ..., y_{iN}).
\]

Using the Peano curve, each object is assigned the value of $x_i$, i.e.
\[
\omega_i \leftrightarrow x_i, \; y(x_i) = (y_{i1}, y_{i2}, ..., y_{iN}).
\]

Consider a test object $\omega^*$, to which the value $x^*$ is assigned. The degree of proximity of the test object $\omega^*$ to one of the objects $\omega_i$ is determined based on the distance on the one-dimensional scale $d_i = |x_i - x^* |$.
Thus, the closest object to the test object has the following property:
\[
d_{min} = \min \{ d_i: 1 \leq i \leq S \}.
\]

The operation of the algorithm to determine the nearest neighbour of the test object involves two phases: data preparation and the search itself. Data preparation consists of calculating the correspondence $\omega_i \rightarrow x_i$ for each input object, including the test object, and ordering the objects by the value of $x$.
\[
x^1(\omega^1) \leq x^2(\omega^2) \leq ... \leq x^S(\omega^S).
\]

@@

It follows from formulae (\ref{hypercubeD0}) and (\ref{maps}) that if an interval $d(M,v)$ is mapped onto the subcube $D_0(M,v)$, then there exists a family of subcubes
\[
D_l(M,v) = y^l(d(M,v)), 1 \leq l \leq L,
\]
connected with the corresponding subintervals $d_l(M,v_l) \subset [0,1]$, where $d(M,v) = d_0(M,v_0), v_0 = v$, such that 
\[
D_l(M,v) = y^l(d_l(M,v)), 1 \leq l \leq L.
\]

Since the hypercube $D$ from (\ref{D}) belongs to the common part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Fig.~\ref{fig2}), then, by introducing an additional function 
\[
g_0(y) = \max \{ |y_i|-2^{-1}: \; 1 \leq i \leq N \},
\] 
the original hypercube $D$ can be represented as
\[
D = \{y^l(x): x \in [0,1], \; g_0(y^l(x)) \leq 0 \}, \; 0 \leq l \leq L,
\] 
i.e. $g_0(y) \leq 0$, if $y \in D$, and $g_0(y) > 0$ if otherwise. Hence, any point $y \in D$ has its preimage $x^l \in [0,1]$ at every mapping $y^l(x), \; 0 \leq l \leq L$, i.e.
\[
y = y^l(x^l), \; x^l \in [0, 1], \; 0 \leq l \leq L.
\]

@@

\textbf{Theorem 1.} Let $y^*$ be an arbitrary point from the domain $D$ belonging to the interval with endpoints $y^\prime, y^{\prime\prime} \in D$, which differ in the values of the only coordinate, and let
\[
|y_j^\prime - y^{\prime\prime}_j| \leq 2^{-p}, \; y^\prime_i = y^{\prime\prime}_i = y_i^*, \; 1 \leq i \leq N, i \ne j,
\] 
where $p$ is an integer, $1 \leq p \leq L-1$, and $j$ is the number of coordinate whose values for points $y^*, y^\prime, y^{\prime\prime}$ are different. Then there exists at least one correspondence $y^l(x), 0 \leq l \leq L$, and preimages $x^*, x^\prime, x^{\prime\prime} \in [0, 1]$ such that
\[
 y^* = y^l(x^*), y^\prime = y^l(x^\prime), y^{\prime\prime} = y^l(x^{\prime\prime}), \\
\max \{ |x^\prime - x^*|, |x^{\prime\prime} - x^*|, |x^\prime - x^{\prime\prime}| \} \leq 2^{-pN}.
\]

@@

The scheme of implementation of the kNN algorithm using multiple Peano curves will look as follows.
For every object $\omega_i$ a set of its preimages on different one-dimensional scales is assigned 
\[
\omega_i \rightarrow \{ x_i^0, x_i^1, ..., x_i^L \},
\] 
where $l, \; 0 \leq l \leq L$, is the number of the one-dimensional interval.

The degree of proximity of a test object $\omega^*$ to one of the known objects $\omega_i$ can be expressed in terms of the distance
\[
d_i^l = |x_i^l - x^{*l}|
\]
on the $l$-th scale.

Thus, the distance $d(\omega_i,\omega^*)$ between two objects $\omega_i$ and $\omega^*$ on a one-dimensional scale will correspond to the minimum distance among all the mappings 
\[
d(\omega_i,\omega^*) = \min_{0\leq l\leq L} d_i^l = \min_{0\leq l\leq L}  |x_i^l - x^{*l}| .
\]
If we find the minimum distance
\[
d_{min} = \min_{1\leq i\leq n} d(\omega_i,\omega^*) ,
\]
we get the nearest neighbour.
